---
title: "Machakos Case Study"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r set_up, echo = FALSE, message = FALSE, warnings = FALSE}

library(leaflet)
library(readr)
library(sf)
library(tidyverse)

shape = read_sf(dsn = "shape/", layer = "subcounties_age")
shape =
  shape %>% 
  rename(pop0_5 = joined0.5,
         pop6_14 = joined6.14,
         pop15_18 = joined15.1,
         pop19_24 = joined19.2,
         pop25_45 = joined25.4,
         pop46_60 = joined46.6,
         pop61_ = joined61.,
         pop_total = joinedTota)

shape$pop0_5 = as.numeric(shape$pop0_5)
shape$pop6_14 = as.numeric(shape$pop6_14)
shape$pop15_18 = as.numeric(shape$pop15_18)
shape$pop19_24 = as.numeric(shape$pop19_24)
shape$pop25_45 = as.numeric(shape$pop25_45)
shape$pop46_60 = as.numeric(shape$pop46_60)
shape$pop61_ = as.numeric(shape$pop61_)
shape$pop_total = as.numeric(shape$pop_total)

facilities_df = read_csv("final_facilities_tidy_3.12.21.csv")

factpal = colorFactor(hcl.colors(4, palette = "Blue-Red 3", alpha = 1), facilities_df$level)
factpal_2 = colorFactor(hcl.colors(3, palette = "Viridis", alpha = 1), facilities_df$ownership)

##labels and pals

pal = colorBin("Blues", domain = shape$pop_total, bins = 5)

labels = glue::glue("<strong>{shape$subcounty}</strong><br/><strong>Total Population:</strong> {shape$pop_total}") %>%
            purrr::map(htmltools::HTML)

pal2 = colorBin("Blues", domain = shape$pop0_5, bins = 5)

labels2 = glue::glue("<strong>{shape$subcounty}</strong><br/><strong>Proportion of Population: </strong>{round(shape$pop0_5/shape$pop_total, 2)*100}%<br/><strong>Population (Ages 0-5):</strong> {shape$pop0_5}<br/><strong>Total Population (All Ages):</strong> {shape$pop_total}") %>%
            purrr::map(htmltools::HTML)

pal3 = colorBin("Blues", domain = shape$pop6_14, bins = 5)

labels3 = glue::glue("<strong>{shape$subcounty}</strong><br/><strong>Proportion of Population: </strong>{round(shape$pop6_14/shape$pop_total, 2)*100}%<br/><strong>Population (Ages 6-14):</strong> {shape$pop6_14}<br/><strong>Total Population (All Ages):</strong> {shape$pop_total}") %>%
            purrr::map(htmltools::HTML)

pal4 = colorBin("Blues", domain = shape$pop15_18, bins = 5)

labels4 = glue::glue("<strong>{shape$subcounty}</strong><br/><strong>Proportion of Population: </strong>{round(shape$pop15_18/shape$pop_total, 2)*100}%<br/><strong>Population (Ages 15-18):</strong> {shape$pop15_18}<br/><strong>Total Population (All Ages):</strong> {shape$pop_total}") %>%
            purrr::map(htmltools::HTML)

pal5 = colorBin("Blues", domain = shape$pop19_24, bins = 5)

labels5 = glue::glue("<strong>{shape$subcounty}</strong><br/><strong>Proportion of Population: </strong>{round(shape$pop19_24/shape$pop_total, 2)*100}%<br/><strong>Population (Ages 19-24):</strong> {shape$pop19_24}<br/><strong>Total Population (All Ages):</strong> {shape$pop_total}") %>%
            purrr::map(htmltools::HTML)

pal6 = colorBin("Blues", domain = shape$pop25_45, bins = 5)

labels6 = glue::glue("<strong>{shape$subcounty}</strong><br/><strong>Proportion of Population: </strong>{round(shape$pop25_45/shape$pop_total, 2)*100}%<br/><strong>Population (Ages 25-45):</strong> {shape$pop25_45}<br/><strong>Total Population (All Ages):</strong> {shape$pop_total}") %>%
            purrr::map(htmltools::HTML)

pal7 = colorBin("Blues", domain = shape$pop46_60, bins = 5)

labels7 = glue::glue("<strong>{shape$subcounty}</strong><br/><strong>Proportion of Population: </strong>{round(shape$pop46_60/shape$pop_total, 2)*100}%<br/><strong>Population (Ages 46-60):</strong> {shape$pop46_60}<br/><strong>Total Population (All Ages):</strong> {shape$pop_total}") %>%
            purrr::map(htmltools::HTML)

pal8 = colorBin("Blues", domain = shape$pop61_, bins = 5)

labels8 = glue::glue("<strong>{shape$subcounty}</strong><br/><strong>Proportion of Population: </strong>{round(shape$pop61_/shape$pop_total, 2)*100}%<br/><strong>Population (Ages 61+):</strong> {shape$pop61_}<br/><strong>Total Population (All Ages):</strong> {shape$pop_total}") %>%
            purrr::map(htmltools::HTML)

```

Population by Age {.tabset .tabset-fade}
-----------------------------------------------------------------------

The following maps show the distribution of the population by age group. Click on the tab for the age group you are interested in viewing. For each map, you can hover over the the map with your mouse to see more details by subcounty.

### Total Population

```{r total, echo=FALSE}
total_map =
  shape %>%
  leaflet() %>%
  addTiles() %>% 
  addPolygons(
     fillColor = ~pal(pop_total),
     weight = 2,
     opacity = 1,
     color = "black",
     dashArray = "1",
     fillOpacity = 0.5,
     highlightOptions = highlightOptions(
     weight = 5,
     color = "#666",
     dashArray = "",
     fillOpacity = 0.7,
      bringToFront = TRUE),
     label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")) %>% 
     addLegend("topright", pal = pal, values = shape$pop_total,
                      title = "Population", opacity = 0.5) %>% 
     addScaleBar(position = "bottomright", options = scaleBarOptions(imperial = FALSE))

total_map
```

### Ages 0-5

```{r 0_5, echo=FALSE}
pop0_5_map =
  shape %>%
  leaflet() %>%
  addTiles() %>% 
  addPolygons(
     fillColor = ~pal2(pop0_5),
     weight = 2,
     opacity = 1,
     color = "black",
     dashArray = "1",
     fillOpacity = 0.5,
     highlightOptions = highlightOptions(
     weight = 5,
     color = "#666",
     dashArray = "",
     fillOpacity = 0.7,
      bringToFront = TRUE),
     label = labels2,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")) %>% 
     addLegend("topright", pal = pal2, values = shape$pop0_5,
                      title = "Population, Ages 0-5", opacity = 0.5) %>% 
     addScaleBar(position = "bottomright", options = scaleBarOptions(imperial = FALSE))

pop0_5_map
```

### Ages 6-14

```{r 6_14, echo=FALSE}
pop6_14_map =
  shape %>%
  leaflet() %>%
  addTiles() %>% 
  addPolygons(
     fillColor = ~pal3(pop6_14),
     weight = 2,
     opacity = 1,
     color = "black",
     dashArray = "1",
     fillOpacity = 0.5,
     highlightOptions = highlightOptions(
     weight = 5,
     color = "#666",
     dashArray = "",
     fillOpacity = 0.7,
      bringToFront = TRUE),
     label = labels3,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")) %>% 
     addLegend("topright", pal = pal3, values = shape$pop6_14,
                      title = "Population, Ages 6-14", opacity = 0.5) %>% 
     addScaleBar(position = "bottomright", options = scaleBarOptions(imperial = FALSE))

pop6_14_map
```

### Ages 15-18

```{r 15_18, echo=FALSE}
pop15_18_map =
  shape %>%
  leaflet() %>%
  addTiles() %>% 
  addPolygons(
     fillColor = ~pal4(pop15_18),
     weight = 2,
     opacity = 1,
     color = "black",
     dashArray = "1",
     fillOpacity = 0.5,
     highlightOptions = highlightOptions(
     weight = 5,
     color = "#666",
     dashArray = "",
     fillOpacity = 0.7,
      bringToFront = TRUE),
     label = labels4,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")) %>% 
     addLegend("topright", pal = pal4, values = shape$pop15_18,
                      title = "Population, Ages 15-18", opacity = 0.5) %>% 
     addScaleBar(position = "bottomright", options = scaleBarOptions(imperial = FALSE))

pop15_18_map
```

### Ages 19-24

```{r 19_24, echo=FALSE}
pop19_24_map =
  shape %>%
  leaflet() %>%
  addTiles() %>% 
  addPolygons(
     fillColor = ~pal5(pop19_24),
     weight = 2,
     opacity = 1,
     color = "black",
     dashArray = "1",
     fillOpacity = 0.5,
     highlightOptions = highlightOptions(
     weight = 5,
     color = "#666",
     dashArray = "",
     fillOpacity = 0.7,
      bringToFront = TRUE),
     label = labels5,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")) %>% 
     addLegend("topright", pal = pal5, values = shape$pop19_24,
                      title = "Population, Ages 19-24", opacity = 0.5) %>% 
     addScaleBar(position = "bottomright", options = scaleBarOptions(imperial = FALSE))

pop19_24_map
```

### Ages 25-45

```{r 25_45, echo=FALSE}
pop25_45_map =
  shape %>%
  leaflet() %>%
  addTiles() %>% 
  addPolygons(
     fillColor = ~pal6(pop25_45),
     weight = 2,
     opacity = 1,
     color = "black",
     dashArray = "1",
     fillOpacity = 0.5,
     highlightOptions = highlightOptions(
     weight = 5,
     color = "#666",
     dashArray = "",
     fillOpacity = 0.7,
      bringToFront = TRUE),
     label = labels6,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")) %>% 
     addLegend("topright", pal = pal6, values = shape$pop25_45,
                      title = "Population, Ages 25-45", opacity = 0.5) %>% 
     addScaleBar(position = "bottomright", options = scaleBarOptions(imperial = FALSE))

pop25_45_map
```

### Ages 46-60

```{r 46_60, echo=FALSE}
pop46_60_map =
  shape %>%
  leaflet() %>%
  addTiles() %>% 
  addPolygons(
     fillColor = ~pal7(pop46_60),
     weight = 2,
     opacity = 1,
     color = "black",
     dashArray = "1",
     fillOpacity = 0.5,
     highlightOptions = highlightOptions(
     weight = 5,
     color = "#666",
     dashArray = "",
     fillOpacity = 0.7,
      bringToFront = TRUE),
     label = labels7,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")) %>% 
     addLegend("topright", pal = pal7, values = shape$pop46_60,
                      title = "Population, Ages 46-60", opacity = 0.5) %>% 
     addScaleBar(position = "bottomright", options = scaleBarOptions(imperial = FALSE))

pop46_60_map
```

### Ages 61+

```{r 61+, echo=FALSE}
pop61_map =
  shape %>%
  leaflet() %>%
  addTiles() %>% 
  addPolygons(
     fillColor = ~pal8(pop61_),
     weight = 2,
     opacity = 1,
     color = "black",
     dashArray = "1",
     fillOpacity = 0.5,
     highlightOptions = highlightOptions(
     weight = 5,
     color = "#666",
     dashArray = "",
     fillOpacity = 0.7,
      bringToFront = TRUE),
     label = labels8,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")) %>% 
     addLegend("topright", pal = pal8, values = shape$pop61_,
                      title = "Population, Ages 61+", opacity = 0.5) %>% 
     addScaleBar(position = "bottomright", options = scaleBarOptions(imperial = FALSE))

pop61_map
```

Healthcare Facilities by Level
-----------------------------------------------------------------------

This map presents the distribution of healthcare facilities by level. Click on the circle markers to see details about healthcare facilities in Machakos County.

```{r facilities_level, echo=FALSE}
leaflet() %>% 
            addProviderTiles(providers$Esri.WorldTopoMap) %>%
            addPolygons(data = shape,
                        weight = 1,
                        opacity = 1,
                        color = "black",
                        dashArray = "1",
                        fillOpacity = 0) %>% 
            addCircleMarkers(data = facilities_df,
                       lat = ~lat,
                       lng = ~lng,
                       radius = ~dplyr::case_when(
                                    facilities_df$level == "Level 5" ~ 15,
                                    facilities_df$level == "Level 4" ~ 10,
                                    TRUE ~ 5),
                       popup = paste(
                           "<strong>", "Facility Name:", "</strong>", facilities_df$facility_name, "<br>", "<strong>", "Ownership:", "</strong>", facilities_df$ownership, "<br>", "<strong>", "Level:", "</strong>", facilities_df$level),
                       color = ~factpal(level), fillOpacity = 0.8, weight = 0.8) %>% 
            addLegend("topright", pal = factpal, values = facilities_df$level,
                      title = "Facility Level", opacity = 0.8) %>% 
            addScaleBar(position = "bottomright", options = scaleBarOptions(imperial = FALSE))
```

Healthcare Facilities by Ownership
-----------------------------------------------------------------------

This map presents the distribution of healthcare facilities by ownership. Click on the circle markers to see details about healthcare facilities in Machakos County.

```{r facilities_ownership, echo=FALSE}
leaflet() %>% 
            addProviderTiles(providers$Esri.WorldTopoMap) %>%
            addPolygons(data = shape,
                        weight = 1,
                        opacity = 1,
                        color = "black",
                        dashArray = "1",
                        fillOpacity = 0) %>% 
            addCircleMarkers(data = facilities_df,
                       lat = ~lat,
                       lng = ~lng,
                       radius = 5,
                       popup = paste(
                           "<strong>", "Facility Name:", "</strong>", facilities_df$facility_name, "<br>", "<strong>", "Ownership:", "</strong>", facilities_df$ownership, "<br>", "<strong>", "Level:", "</strong>", facilities_df$level),
                       color = ~factpal_2(ownership), fillOpacity = 0.8, weight = 0.8) %>% 
            addLegend("topright", pal = factpal_2, values = facilities_df$ownership,
                      title = "Facility Ownership", opacity = 0.8) %>% 
            addScaleBar(position = "bottomright", options = scaleBarOptions(imperial = FALSE))
```
